---
title: 'STA 444 Practical F24 (MWF)'
author: "Mandi Bluth"
date: "`r format(Sys.time(),  '%B %d, %Y')`"
output: pdf_document
---

# Instructions

## You may use the book during the exam and may take until the end of the period.

* **Change the header information within the RMD to contain your own name.**

* **Answer all exercise prompts within the RMD.  All code must be shown.**

* **Place answers into the blank R chunks given for each required response.**

* **Compile the RMD into a PDF when finished.**

* **Ensure all code is visible within the PDF.**

* **Submit the PDF through our Canvas portal.**

* **The practical is not graded - try your best!**

## Exercise 1

Specifically load the packages `ggplot2`, `dplyr`, and `broom`.  Be sure the code is displayed.  It is okay if warnings or messages are output.

```{r}
library(ggplot2)
library(dplyr)
library(broom)
```


## Exercise 2

Load the data set `mtcars` from base R.  Display the first 6 rows of this data frame.

```{r}
data(mtcars)
head(mtcars)
```


## Exercise 3

Create a new column within `mtcars` called `wt01` that takes on the value `LightWeight` if less than the mean weight (`wt`) of the data set, or takes on the value `HeavyWeight` if greater than or equal to the mean weight. *This question should use dplyr commands to achieve the result.*

```{r}
mtcars <- mtcars %>%
    mutate(wt01 = ifelse(wt < mean(mtcars$wt), 'LightWeight', 'HeavyWeight'))
head(mtcars)
```

## Exercise 4

Using the `mtcars` data, create a scatter graph with `hp` on the x-axis and `mpg` on the y-axis. Color the graph using the `cyl`. To get unique groups add the command `factor(cyl)` rather than just the variable `cyl` when coloring your graph.

```{r}
ggplot(data=mtcars,
       mapping=aes(x=hp, y=mpg, color=factor(cyl)))+
  geom_point()
```

## Exercise 5

Split your graph above into a two paneled graph by adding a facet for `wt01`.

```{r}
ggplot(data=mtcars,
       mapping=aes(x=hp, y=mpg, color=factor(cyl)))+
  geom_point()+
  facet_grid(cols=vars(wt01))
```

## Exercise 6

Using the `mtcars` data, calculate the mean and standard deviation of `mpg` for each `cyl` and `wt01` combination. *Hint: group_by().*

```{r}
mtcars %>% group_by(cyl, wt01) %>% summarise( mean(mpg), sd(mpg))
```

## Exercise 7

Estimate the model below using the `mtcars` data.
```{r}
model.mtcars <- lm(data = mtcars, mpg ~ hp*wt01)
```

Display the resulting coefficient estimates using the `broom` package.  Be sure to include the confidence intervals for each parameter.

```{r}
broom::tidy(model.mtcars, conf.int=TRUE)
```

